Setup
```{r}
rm(list=ls())
library(arm)
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
```

Load data
```{r}
data<-read.csv("data/W8QDEB2.csv",header = TRUE, stringsAsFactors=TRUE)
head(data)
summary(data)
dim(data)
```
continuous variables: "W1GrssyrHH", "W1yschat1", "W2ghq12scr", "W4schatYP", "W6DebtattYP", "W8DGHQSC", "W8QDEB2"

Exploratory analysis of categorical variables
#"W1wrk1aMP" vs "W8QDEB2"
```{r}
W1wrk1aMP<-ggplot(data=data, aes(x=factor(W1wrk1aMP),y=W8QDEB2)) +
  geom_boxplot()
  
W1wrk1aMP
```
#"W1hea2MP" vs "W8QDEB2"
```{r}
W1hea2MP<-ggplot(data=data, aes(x=factor(W1hea2MP),y=W8QDEB2)) +
  geom_boxplot()
  
W1hea2MP
```

#"W1NoldBroHS" vs "W8QDEB2"
```{r}
W1NoldBroHS<-ggplot(data=data, aes(x=factor(W1NoldBroHS),y=W8QDEB2)) +
  geom_boxplot()
  
W1NoldBroHS
```

#"W1InCarHH" vs "W8QDEB2"
```{r}
W1InCarHH<-ggplot(data=data, aes(x=factor(W1InCarHH),y=W8QDEB2)) +
  geom_boxplot()
  
W1InCarHH
```

#"W1hous12HH" vs "W8QDEB2"
```{r}
W1hous12HH<-ggplot(data=data, aes(x=factor(W1hous12HH),y=W8QDEB2)) +
  geom_boxplot()
  
W1hous12HH
```

#"W1usevcHH" vs "W8QDEB2"
```{r}
W1usevcHH<-ggplot(data=data, aes(x=factor(W1usevcHH),y=W8QDEB2)) +
  geom_boxplot()
  
W1usevcHH
```

#"W1hiqualmum" vs "W8QDEB2"
```{r}
W1hiqualmum<-ggplot(data=data, aes(x=factor(W1hiqualmum),y=W8QDEB2)) +
  geom_boxplot()
  
W1hiqualmum
```

#"W1hiqualdad" vs "W8QDEB2"
```{r}
W1hiqualdad<-ggplot(data=data, aes(x=factor(W1hiqualdad),y=W8QDEB2)) +
  geom_boxplot()
  
W1hiqualdad
```

#"W1wrkfulldad" vs "W8QDEB2"
```{r}
W1wrkfulldad<-ggplot(data=data, aes(x=factor(W1wrkfulldad),y=W8QDEB2)) +
  geom_boxplot()
  
W1wrkfulldad
```

#"W1wrkfullmum" vs "W8QDEB2"
```{r}
W1wrkfullmum<-ggplot(data=data, aes(x=factor(W1wrkfullmum),y=W8QDEB2)) +
  geom_boxplot()
  
W1wrkfullmum
```

#"W1empsmum" vs "W8QDEB2"
```{r}
W1empsmum<-ggplot(data=data, aes(x=factor(W1empsmum),y=W8QDEB2)) +
  geom_boxplot()
  
W1empsmum
```

#"W1empsdad" vs "W8QDEB2"
```{r}
W1empsdad<-ggplot(data=data, aes(x=factor(W1empsdad),y=W8QDEB2)) +
  geom_boxplot()
  
W1empsdad
```

#"IndSchool" vs "W8QDEB2"
```{r}
IndSchool<-ggplot(data=data, aes(x=factor(IndSchool),y=W8QDEB2)) +
  geom_boxplot()
  
IndSchool
```

#"W1marstatmum" vs "W8QDEB2"
```{r}
W1marstatmum<-ggplot(data=data, aes(x=factor(W1marstatmum),y=W8QDEB2)) +
  geom_boxplot()
  
W1marstatmum
```

#"W1depkids" vs "W8QDEB2"
```{r}
W1depkids<-ggplot(data=data, aes(x=factor(W1depkids),y=W8QDEB2)) +
  geom_boxplot()
  
W1depkids
```
#"W1famtyp2" vs "W8QDEB2"
```{r}
W1famtyp2<-ggplot(data=data, aes(x=factor(W1famtyp2),y=W8QDEB2)) +
  geom_boxplot()
  
W1famtyp2
```

#"W1nssecfam" vs "W8QDEB2"
```{r}
W1nssecfam<-ggplot(data=data, aes(x=factor(W1nssecfam),y=W8QDEB2)) +
  geom_boxplot()
  
W1nssecfam
```

#"W1ethgrpYP" vs "W8QDEB2"
```{r}
W1ethgrpYP<-ggplot(data=data, aes(x=factor(W1ethgrpYP),y=W8QDEB2)) +
  geom_boxplot()
  
W1ethgrpYP
```

#"W1heposs9YP" vs "W8QDEB2"
```{r}
W1heposs9YP<-ggplot(data=data, aes(x=factor(W1heposs9YP),y=W8QDEB2)) +
  geom_boxplot()
  
W1heposs9YP
```

#"W1hwndayYP" vs "W8QDEB2"
```{r}
W1hwndayYP<-ggplot(data=data, aes(x=factor(W1hwndayYP),y=W8QDEB2)) +
  geom_boxplot()
  
W1hwndayYP
```

#"W1truantYP" vs "W8QDEB2"
```{r}
W1truantYP<-ggplot(data=data, aes(x=factor(W1truantYP),y=W8QDEB2)) +
  geom_boxplot()
  
W1truantYP
```

#"W1alceverYP" vs "W8QDEB2"
```{r}
W1alceverYP<-ggplot(data=data, aes(x=factor(W1alceverYP),y=W8QDEB2)) +
  geom_boxplot()
  
W1alceverYP
```

#"W1bulrc" vs "W8QDEB2"
```{r}
W1bulrc<-ggplot(data=data, aes(x=factor(W1bulrc),y=W8QDEB2)) +
  geom_boxplot()
  
W1bulrc
```

#"W1disabYP" vs "W8QDEB2"
```{r}
W1disabYP<-ggplot(data=data, aes(x=factor(W1disabYP),y=W8QDEB2)) +
  geom_boxplot()
  
W1disabYP
```

Exploratory analysis of continuous variables

#"W1GrssyrHH" vs "W8QDEB2"
```{r}
W1GrssyrHH<-ggplot(data, aes(x = W1GrssyrHH, y = W8QDEB2)) + 
  geom_point(alpha=0.5)

W1GrssyrHH
```

#"W1yschat1" vs "W8QDEB2"
```{r}
W1yschat1<-ggplot(data, aes(x = W1yschat1, y = W8QDEB2)) + 
  geom_point(alpha=0.5)

W1yschat1
```

W2ghq12scr<-
W4schatYP<-
W6DebtattYP<-
W8DGHQSC<-

Calculate percentage of missing vlaues among predictors
```{r}
percentages <- c()
for (col in names(data)) {
    column <- data[[col]]
    count <- sum(column >= -999 & column <= -1)
    total_values <- length(column)
    percentage <- (count / total_values) * 100
    percentages <- c(percentages, percentage)
}

missing <- data.frame(Column = names(data), Percentage = percentages)
missing
```


Categorical variables with missing values

# remove data with less than 10% missing values
```{r}
new.data <- subset(data, W1wrk1aMP>-1 &
                         W1condur5MP>-1 &
                         W1hea2MP>-1&
                         W1NoldBroHS>-1&
                         W1InCarHH>-1&
                         W1hous12HH>-1&
                         W1usevcHH>-1&
                         W1hiqualmum>-1&
                         W1wrkfullmum>-1&
                         W1empsmum>-1&
                         W1marstatmum>-1&
                         W1depkids>-1&
                         W1famtyp2>-1&
                         W1nssecfam>-1&
                         W1ethgrpYP>-1&
                         W1heposs9YP>-1&			
                         W1hwndayYP>-1&			
                         W1truantYP>-1&			
                         W1alceverYP>-1&		
                         W1bulrc>-1&		
                         W1disabYP>-1&		
                         W2disc1YP>-1&			
                         W2depressYP>-1&
                         W4CannTryYP>-1&			
                         W4NamesYP>-1&			
                         W4RacismYP>-1&			
                         W4empsYP>-1&			
                         W5JobYP>-1&			
                         W5EducYP>-1&			
                         W5Apprent1YP>-1&
                         W6JobYP>-1&			
                         W6UnivYP>-1&
                         W6OwnchiDV>-1&
                         W8DMARSTAT>-1&			
                         W8TENURE
                         )
dim(new.data)
```

# replace data with NA for more than 10% missing values
```{r}
data[, columns_more_than_10_percent] <- lapply(data[, columns_more_than_10_percent], function(x) {
    x[x >= -999 & x <= -1] <- NA
    return(x)
})
```

Continuous variables with missing values
```{r}
```