Setup
```{r}
rm(list=ls())
library(arm)
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
```

Load data
```{r}
data<-read.csv("../data/W8QDEB2.csv",header = TRUE, stringsAsFactors=TRUE)
head(data)
summary(data)
dim(data)
```

Calculate percentage of missing values among predictors

```{r}
percentages <- c()
for (col in names(data)) {
    column <- data[[col]]
    count <- sum(column >= -999 & column <= -1)
    total_values <- length(column)
    percentage <- (count / total_values) * 100
    percentages <- c(percentages, percentage)
}

missing <- data.frame(Column = names(data), Percentage = percentages)
missing
```
# Summary 

Continuous: (more than 30 percent missing)
  "W1GrssyrHH": 46.56 (remove from final model to see if regression changes substantially)
  
  
Categorical: (more than 10 percent missing)
  "W1hiqualdad": 26.41
  "W1wrkfulldad": 20.60
  "W1empsdad": 20.60
  "W4AlcFreqYP": 17.41
  "W4Childck1YP":99.44
  "W6EducYP":38.99
  "W6Apprent1YP":38.99
  "W6Childliv":97.22

# Clean data

Remove data with less than 10% missing values for categorical variables

```{r}
new.data <- subset(data, W1wrk1aMP>-1 &
                         W1condur5MP>-1 &
                         W1hea2MP>-1&
                         W1NoldBroHS>-1&
                         W1InCarHH>-1&
                         W1hous12HH>-1&
                         W1usevcHH>-1&
                         W1hiqualmum>-1&
                         W1wrkfullmum>-1&
                         W1empsmum>-1&
                         W1marstatmum>-1&
                         W1depkids>-1&
                         W1famtyp2>-1&
                         W1nssecfam>-1&
                         W1ethgrpYP>-1&
                         W1heposs9YP>-1&			
                         W1hwndayYP>-1&			
                         W1truantYP>-1&			
                         W1alceverYP>-1&		
                         W1bulrc>-1&		
                         W1disabYP>-1&		
                         W2disc1YP>-1&			
                         W2depressYP>-1&
                         W4CannTryYP>-1&			
                         W4NamesYP>-1&			
                         W4RacismYP>-1&			
                         W4empsYP>-1&			
                         W5JobYP>-1&			
                         W5EducYP>-1&			
                         W5Apprent1YP>-1&
                         W6JobYP>-1&			
                         W6UnivYP>-1&
                         W6OwnchiDV>-1&
                         W8DMARSTAT>-1&			
                         W8TENURE
                         )
dim(new.data)
```

Replace data with NA

```{r}
#change all the negative values to NA
new.data.1 <- new.data %>%
  mutate(across(everything(), ~replace(., . < 0, NA)))
```

Create a "missing level" for categorical variables

```{r}
create_missing_category <- function(data, var_name) {
  # Replace NA values with a new level called "Missing"
  data[[var_name]][is.na(data[[var_name]])] <- "Missing"
  
  # Convert the variable to a factor with the new level
  data[[var_name]] <- factor(data[[var_name]])
  
  return(data)
}
```

```{r}
new.data.2 <- create_missing_category(new.data.1, "W1wrkfulldad")
new.data.2 <- create_missing_category(new.data.2, "W4AlcFreqYP")
new.data.2 <- create_missing_category(new.data.2, "W4Childck1YP")
new.data.2 <- create_missing_category(new.data.2, "W6EducYP")
new.data.2 <- create_missing_category(new.data.2, "W8TENURE")
new.data.2 <- create_missing_category(new.data.2, "W1empsdad")
new.data.2 <- create_missing_category(new.data.2, "W6Apprent1YP")
new.data.2 <- create_missing_category(new.data.2, "W6Childliv")
new.data.2 <- create_missing_category(new.data.2, "W1hiqualdad")
```

Check the missing values in new.data

```{r}
colSums(is.na(new.data.2))
```

Store the cleaned data in a csv file

```{r}
write.csv(new.data.2, file = "../data/W8QDEB2_cleaned.csv")
```