Setup
```{r}
rm(list=ls())
library(arm)
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
```

Load data
```{r}
data<-read.csv("../data/W8QDEB2_cleaned.csv",header = TRUE, stringsAsFactors=TRUE)
head(data)
summary(data)
dim(data)
```
Continuous variables: "W1GrssyrHH", "W1yschat1", "W2ghq12scr", "W4schatYP", "W6DebtattYP", "W8DGHQSC"

Binary variables: "W1hea2MP", "W1usevcHH", "IndSchool", "W1truantYP", "W1alceverYP", "W1bulrc", "W5JobYP", "W5EducYP", "W5Apprent1YP", "W6JobYP", "W6UnivYP", "W6EducYP", "W6Apprent1YP", "W6OwnchiDV", "W6Childliv"

# Exploratory analysis of binary variables

"W1hea2MP" vs "W8QDEB2"
```{r}
W1hea2MP<-ggplot(data=data, aes(x=factor(W1hea2MP),y=W8QDEB2)) +
  geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
  
W1hea2MP
```
```{r}
summary(lm(W8QDEB2 ~ factor(W1hea2MP), data=data))
```
not significant

"W1usevcHH" vs "W8QDEB2"
```{r}
W1usevcHH<-ggplot(data=data, aes(x=factor(W1usevcHH),y=W8QDEB2)) +
  geom_boxplot()
  
W1usevcHH
```
```{r}
summary(lm(W8QDEB2 ~ factor(W1usevcHH), data=data))
```
 significant at 1 percent. 

"IndSchool" vs "W8QDEB2"
```{r}
IndSchool<-ggplot(data=data, aes(x=factor(IndSchool),y=W8QDEB2)) +
  geom_boxplot()
  
IndSchool
```
```{r}
summary(lm(W8QDEB2 ~ factor(IndSchool), data=data))
```
highly significant

"W1truantYP" vs "W8QDEB2"
```{r}
W1truantYP<-ggplot(data=data, aes(x=factor(W1truantYP),y=W8QDEB2)) +
  geom_boxplot()
  
W1truantYP
```

```{r}
summary(lm(W8QDEB2 ~ factor(W1truantYP), data=data))
```
not significant

"W1alceverYP" vs "W8QDEB2"
```{r}
W1alceverYP<-ggplot(data=data, aes(x=factor(W1alceverYP),y=W8QDEB2)) +
  geom_boxplot()
  
W1alceverYP
```

```{r}
summary(lm(W8QDEB2 ~ factor(W1alceverYP), data=data))
```
not significant

"W1bulrc" vs "W8QDEB2"
```{r}
W1bulrc<-ggplot(data=data, aes(x=factor(W1bulrc),y=W8QDEB2)) +
  geom_boxplot()
  
W1bulrc
```

```{r}
summary(lm(W8QDEB2 ~ factor(W1bulrc), data=data))
```
not significant

W5JobYP vs W8QDEB2
```{r}
W5JobYP <- ggplot(data=data, aes(x=factor(W5JobYP), y=W8QDEB2)) + geom_boxplot()
W5JobYP
```

```{r}
summary(lm(W8QDEB2 ~ factor(W5JobYP), data=data))
```
not significant

W5EducYP vs W8QDEB2
```{r}
W5EducYP <- ggplot(data=data, aes(x=factor(W5EducYP), y=W8QDEB2)) + geom_boxplot()
W5EducYP
```

```{r}
summary(lm(W8QDEB2 ~ factor(W5EducYP), data=data))
```
not significant

W5Apprent1YP vs W8QDEB2
```{r}
W5Apprent1YP <- ggplot(data=data, aes(x=factor(W5Apprent1YP), y=W8QDEB2)) + geom_boxplot()
W5Apprent1YP
```

```{r}
summary(lm(W8QDEB2 ~ factor(W5Apprent1YP), data=data))
```
not significant

W6JobYP vs W8QDEB2
```{r}
W6JobYP <- ggplot(data=data, aes(x=factor(W6JobYP), y=W8QDEB2)) + geom_boxplot()
W6JobYP
```

```{r}
summary(lm(W8QDEB2 ~ factor(W6JobYP), data=data))
```
not significant

W6UnivYP vs W8QDEB2
```{r}
W6UnivYP <- ggplot(data=data, aes(x=factor(W6UnivYP), y=W8QDEB2)) + geom_boxplot()
W6UnivYP
```

```{r}
summary(lm(W8QDEB2 ~ factor(W6UnivYP), data=data))
```
significant

W6EducYP vs W8QDEB2
```{r}
W6EducYP <- ggplot(data=data, aes(x=factor(W6EducYP), y=W8QDEB2)) + geom_boxplot()
W6EducYP
```

```{r}
summary(lm(W8QDEB2 ~ factor(W6EducYP), data=data))
```
missing level significant at 1 percent. 

W6Apprent1YP vs W8QDEB2
```{r}
W6Apprent1YP <- ggplot(data=data, aes(x=factor(W6Apprent1YP), y=W8QDEB2)) + geom_boxplot()
W6Apprent1YP
```

```{r}
summary(lm(W8QDEB2 ~ factor(W6Apprent1YP), data=data))
```
missing level significant at 5%

W6OwnchiDV vs W8QDEB2
```{r}
W6OwnchiDV <- ggplot(data=data, aes(x=factor(W6OwnchiDV), y=W8QDEB2)) + geom_boxplot()
W6OwnchiDV
```

```{r}
summary(lm(W8QDEB2 ~ factor(W6OwnchiDV), data=data))
```
not significant

W6Childliv vs W8QDEB2
```{r}
W6Childliv <- ggplot(data=data, aes(x=factor(W6Childliv), y=W8QDEB2)) + geom_boxplot()
W6Childliv
```

```{r}
summary(lm(W8QDEB2 ~ factor(W6Childliv), data=data))
```
not significant

# Exploratory analysis of categorical variables

"W1wrk1aMP" vs "W8QDEB2"
```{r}
W1wrk1aMP<-ggplot(data=data, aes(x=factor(W1wrk1aMP),y=W8QDEB2)) +
  geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
  
W1wrk1aMP
```
```{r}
summary(lm(W8QDEB2 ~ factor(W1wrk1aMP), data=data))
```
not significant

"W1NoldBroHS" vs "W8QDEB2"
```{r}
W1NoldBroHS<-ggplot(data=data, aes(x=factor(W1NoldBroHS),y=W8QDEB2)) +
    geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
  
W1NoldBroHS
```

```{r}
summary(lm(W8QDEB2 ~ factor(W1NoldBroHS), data=data))
```
not significant


"W1InCarHH" vs "W8QDEB2"
```{r}
W1InCarHH<-ggplot(data=data, aes(x=factor(W1InCarHH),y=W8QDEB2)) +
  geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
  
W1InCarHH
```
```{r}
summary(lm(W8QDEB2 ~ factor(W1InCarHH), data=data))
```
not significant

"W1hous12HH" vs "W8QDEB2"
```{r}
W1hous12HH<-ggplot(data=data, aes(x=factor(W1hous12HH),y=W8QDEB2)) +
  geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
  
W1hous12HH
```
```{r}
summary(lm(W8QDEB2 ~ factor(W1hous12HH), data=data))
```
level 5 is significant at the 5% level

"W1hiqualmum" vs "W8QDEB2"
```{r}
W1hiqualmum<-ggplot(data=data, aes(x=factor(W1hiqualmum),y=W8QDEB2)) +
  geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
  
W1hiqualmum
```

```{r}
summary(lm(W8QDEB2 ~ factor(W1hiqualmum), data=data))
```
can be combined, only level 3 is significant at the 5% level.

"W1hiqualdad" vs "W8QDEB2"
```{r}
W1hiqualdad<-ggplot(data=data, aes(x=factor(W1hiqualdad),y=W8QDEB2)) +
  geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
  
W1hiqualdad
```
```{r}
summary(lm(W8QDEB2 ~ factor(W1hiqualdad), data=data))
```
Level 16, 4, and 9 are significant at the 5% level.

"W1wrkfulldad" vs "W8QDEB2"
```{r}
W1wrkfulldad<-ggplot(data=data, aes(x=factor(W1wrkfulldad),y=W8QDEB2)) + geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
  
W1wrkfulldad
```
```{r}
summary(lm(W8QDEB2 ~ factor(W1wrkfulldad), data=data))
```
not significant

"W1wrkfullmum" vs "W8QDEB2"
```{r}
W1wrkfullmum<-ggplot(data=data, aes(x=factor(W1wrkfullmum),y=W8QDEB2)) +  geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
  
W1wrkfullmum
```
```{r}
summary(lm(W8QDEB2 ~ factor(W1wrkfullmum), data=data))
```
not significant

"W1empsmum" vs "W8QDEB2"
```{r}
W1empsmum<-ggplot(data=data, aes(x=factor(W1empsmum),y=W8QDEB2)) +
  geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
  
W1empsmum
```
```{r}
summary(lm(W8QDEB2 ~ factor(W1empsmum), data=data))
```

"W1empsdad" vs "W8QDEB2"
```{r}
W1empsdad<-ggplot(data=data, aes(x=factor(W1empsdad),y=W8QDEB2)) +
  geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
  
W1empsdad
```
```{r}
summary(lm(W8QDEB2 ~ factor(W1empsdad), data=data))
```
not significant

"W1marstatmum" vs "W8QDEB2"
```{r}
W1marstatmum<-ggplot(data=data, aes(x=factor(W1marstatmum),y=W8QDEB2)) + geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
  
W1marstatmum
```
```{r}
summary(lm(W8QDEB2 ~ factor(W1marstatmum), data=data))
```
level 2 is significant at the 5% level. 

"W1depkids" vs "W8QDEB2"
```{r}
W1depkids<-ggplot(data=data, aes(x=factor(W1depkids),y=W8QDEB2)) +
  geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
  
W1depkids
```
```{r}
summary(lm(W8QDEB2 ~ factor(W1depkids), data=data))
```
not significant

"W1famtyp2" vs "W8QDEB2"
```{r}
W1famtyp2<-ggplot(data=data, aes(x=factor(W1famtyp2),y=W8QDEB2)) +
  geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
  
W1famtyp2
```
```{r}
summary(lm(W8QDEB2 ~ factor(W1famtyp2), data=data))
```
not significant

"W1nssecfam" vs "W8QDEB2"
```{r}
W1nssecfam<-ggplot(data=data, aes(x=factor(W1nssecfam),y=W8QDEB2)) +
  geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
  
W1nssecfam
```
```{r}
summary(lm(W8QDEB2 ~ factor(W1nssecfam), data=data))
```
significant at the 1 percent.

"W1ethgrpYP" vs "W8QDEB2"
```{r}
W1ethgrpYP<-ggplot(data=data, aes(x=factor(W1ethgrpYP),y=W8QDEB2)) +
  geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
  
W1ethgrpYP
```
```{r}
summary(lm(W8QDEB2 ~ factor(W1ethgrpYP), data=data))
```
level 2 and 3 significant at the 5% level.

"W1heposs9YP" vs "W8QDEB2"
```{r}
W1heposs9YP<-ggplot(data=data, aes(x=factor(W1heposs9YP),y=W8QDEB2)) +
  geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
  
W1heposs9YP
```
```{r}
summary(lm(W8QDEB2 ~ factor(W1heposs9YP), data=data))
```
significant at the 1% level.

"W1hwndayYP" vs "W8QDEB2"
```{r}
W1hwndayYP<-ggplot(data=data, aes(x=factor(W1hwndayYP),y=W8QDEB2)) +
  geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
  
W1hwndayYP
```
```{r}
summary(lm(W8QDEB2 ~ factor(W1hwndayYP), data=data))
```
not significant

"W1disabYP" vs "W8QDEB2"
```{r}
W1disabYP<-ggplot(data=data, aes(x=factor(W1disabYP),y=W8QDEB2)) +
  geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
  
W1disabYP
```
```{r}
summary(lm(W8QDEB2 ~ factor(W1disabYP), data=data))
```
not significant

W2disc1YP vs W8QDEB2
```{r}
W2disc1YP <- ggplot(data=data, aes(x=factor(W2disc1YP), y=W8QDEB2)) +
  geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
W2disc1YP
```
```{r}
summary(lm(W8QDEB2 ~ factor(W2disc1YP), data=data))
```
not significant

W2depressYP vs W8QDEB2
```{r}
W2depressYP <- ggplot(data=data, aes(x=factor(W2depressYP), y=W8QDEB2)) + geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
W2depressYP
```
```{r}
summary(lm(W8QDEB2 ~ factor(W2depressYP), data=data))
```
not significant

W4AlcFreqYP vs W8QDEB2
```{r}
W4AlcFreqYP <- ggplot(data=data, aes(x=factor(W4AlcFreqYP), y=W8QDEB2)) + geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
W4AlcFreqYP
```
```{r}
summary(lm(W8QDEB2 ~ factor(W4AlcFreqYP), data=data))
```
not significant

W4CannTryYP vs W8QDEB2
```{r}
W4CannTryYP <- ggplot(data=data, aes(x=factor(W4CannTryYP), y=W8QDEB2)) + geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
W4CannTryYP
```
```{r}
summary(lm(W8QDEB2 ~ factor(W4CannTryYP), data=data))
```
not significant

W4NamesYP vs W8QDEB2
```{r}
W4NamesYP <- ggplot(data=data, aes(x=factor(W4NamesYP), y=W8QDEB2)) + geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
W4NamesYP
```
```{r}
summary(lm(W8QDEB2 ~ factor(W4NamesYP), data=data))
```
not significant

W4RacismYP vs W8QDEB2
```{r}
W4RacismYP <- ggplot(data=data, aes(x=factor(W4RacismYP), y=W8QDEB2)) + geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
W4RacismYP
```
```{r}
summary(lm(W8QDEB2 ~ factor(W4RacismYP), data=data))
```
significant at the 5% level.

W4empsYP vs W8QDEB2
```{r}
W4empsYP <- ggplot(data=data, aes(x=factor(W4empsYP), y=W8QDEB2)) + geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
W4empsYP
```
```{r}
summary(lm(W8QDEB2 ~ factor(W4empsYP), data=data))
```
not significant

W4Childck1YP vs W8QDEB2
```{r}
W4Childck1YP <- ggplot(data=data, aes(x=factor(W4Childck1YP), y=W8QDEB2)) + geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
W4Childck1YP
```
```{r}
summary(lm(W8QDEB2 ~ factor(W4Childck1YP), data=data))
```
not significant

# W6acqno vs W8QDEB2
```{r}
W6acqno <- ggplot(data=data, aes(x=factor(W6acqno), y=W8QDEB2)) + geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
W6acqno
```
```{r}
summary(lm(W8QDEB2 ~ factor(W6acqno), data=data))
```
significant

W6gcse vs W8QDEB2
```{r}
W6gcse <- ggplot(data=data, aes(x=factor(W6gcse), y=W8QDEB2)) + geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
W6gcse
```
```{r}
summary(lm(W8QDEB2 ~ factor(W6gcse), data=data))
```
not significant

W6als vs W8QDEB2
```{r}
W6als <- ggplot(data=data, aes(x=factor(W6als), y=W8QDEB2)) + geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
W6als
```
```{r}
summary(lm(W8QDEB2 ~ factor(W6als), data=data))
```
not significant

W8DMARSTAT vs W8QDEB2
```{r}
W8DMARSTAT <- ggplot(data=data, aes(x=factor(W8DMARSTAT), y=W8QDEB2)) + geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
W8DMARSTAT
```
```{r}
summary(lm(W8QDEB2 ~ factor(W8DMARSTAT), data=data))
```
level 2 significant at the 5 percent level.

W8TENURE vs W8QDEB2
```{r}
W8TENURE <- ggplot(data=data, aes(x=factor(W8TENURE), y=W8QDEB2)) + geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 25000))
W8TENURE
```
```{r}
summary(lm(W8QDEB2 ~ factor(W8TENURE), data=data))
```
level 2 significant. 

W8DACTIVITY vs W8QDEB2
```{r}
W8DACTIVITY <- ggplot(data=data, aes(x=factor(W8DACTIVITY), y=W8QDEB2)) + geom_boxplot(outlier.shape = NA)  +
  coord_cartesian(ylim = c(0, 40000))
W8DACTIVITY
```
```{r}
summary(lm(W8QDEB2 ~ factor(W8DACTIVITY), data=data))
```
level 2 and 6 significant at the 1% level. 

Exploratory analysis of continuous variables

#"W1GrssyrHH" vs "W8QDEB2"
```{r}
W1GrssyrHH<-ggplot(data, aes(x = W1GrssyrHH, y = W8QDEB2)) + 
  geom_point(alpha=0.5)

W1GrssyrHH
```
```{r}
summary(lm(W8QDEB2 ~ W1GrssyrHH, data=data))
```
highly significant

#"W1yschat1" vs "W8QDEB2"
```{r}
W1yschat1<-ggplot(data, aes(x = W1yschat1, y = W8QDEB2)) + 
  geom_point(alpha=0.5)

W1yschat1
```
```{r}
summary(lm(W8QDEB2 ~ W1yschat1, data=data))
```
not significant 

# W2ghq12scr vs W8QDEB2
```{r}
W2ghq12scr <- ggplot(data=data, aes(x=W2ghq12scr, y=W8QDEB2)) +
  geom_point(alpha=0.5)
W2ghq12scr
```
```{r}
summary(lm(W8QDEB2 ~ W2ghq12scr, data=data))
```
not significant

# W4schatYP vs W8QDEB2
```{r}
W4schatYP <- ggplot(data=data, aes(x=W4schatYP, y=W8QDEB2)) + geom_point(alpha=0.5)
W4schatYP
```
```{r}
summary(lm(W8QDEB2 ~ W4schatYP, data=data))
```
not significant

# W6DebtattYP vs W8QDEB2
```{r}
W6DebtattYP <- ggplot(data=data, aes(x=W6DebtattYP, y=W8QDEB2)) + geom_point(alpha=0.5)
W6DebtattYP
```
```{r}
summary(lm(W8QDEB2 ~ W6DebtattYP, data=data))
```
not significant 

# W8DGHQSC vs W8QDEB2
```{r}
W8DGHQSC <- ggplot(data=data, aes(x=W8DGHQSC, y=W8QDEB2)) + geom_point(alpha=0.5)
W8DGHQSC
```

```{r}
summary(lm(W8QDEB2 ~ W8DGHQSC, data=data))
```
not significant 