Setup
```{r}
rm(list=ls())
library(arm)
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
```

Load data
```{r}
data<-read.csv("../data/W8QDEB2.csv",header = TRUE, stringsAsFactors=TRUE)
head(data)
summary(data)
dim(data)
```

Exploratory analysis of categorical variables

#"W1wrk1aMP" vs "W8QDEB2"
```{r}
W1wrk1aMP<-ggplot(data=data, aes(x=factor(W1wrk1aMP),y=W8QDEB2)) +
  geom_boxplot()
  
W1wrk1aMP
```
#"W1hea2MP" vs "W8QDEB2"
```{r}
W1hea2MP<-ggplot(data=data, aes(x=factor(W1hea2MP),y=W8QDEB2)) +
  geom_boxplot()
  
W1hea2MP
```

#"W1NoldBroHS" vs "W8QDEB2"
```{r}
W1NoldBroHS<-ggplot(data=data, aes(x=factor(W1NoldBroHS),y=W8QDEB2)) +
  geom_boxplot()
  
W1NoldBroHS
```

#"W1InCarHH" vs "W8QDEB2"
```{r}
W1InCarHH<-ggplot(data=data, aes(x=factor(W1InCarHH),y=W8QDEB2)) +
  geom_boxplot()
  
W1InCarHH
```

#"W1hous12HH" vs "W8QDEB2"
```{r}
W1hous12HH<-ggplot(data=data, aes(x=factor(W1hous12HH),y=W8QDEB2)) +
  geom_boxplot()
  
W1hous12HH
```

#"W1usevcHH" vs "W8QDEB2"
```{r}
W1usevcHH<-ggplot(data=data, aes(x=factor(W1usevcHH),y=W8QDEB2)) +
  geom_boxplot()
  
W1usevcHH
```

#"W1hiqualmum" vs "W8QDEB2"
```{r}
W1hiqualmum<-ggplot(data=data, aes(x=factor(W1hiqualmum),y=W8QDEB2)) +
  geom_boxplot()
  
W1hiqualmum
```

#"W1hiqualdad" vs "W8QDEB2"
```{r}
W1hiqualdad<-ggplot(data=data, aes(x=factor(W1hiqualdad),y=W8QDEB2)) +
  geom_boxplot()
  
W1hiqualdad
```

#"W1wrkfulldad" vs "W8QDEB2"
```{r}
W1wrkfulldad<-ggplot(data=data, aes(x=factor(W1wrkfulldad),y=W8QDEB2)) +
  geom_boxplot()
  
W1wrkfulldad
```

#"W1wrkfullmum" vs "W8QDEB2"
```{r}
W1wrkfullmum<-ggplot(data=data, aes(x=factor(W1wrkfullmum),y=W8QDEB2)) +
  geom_boxplot()
  
W1wrkfullmum
```

#"W1empsmum" vs "W8QDEB2"
```{r}
W1empsmum<-ggplot(data=data, aes(x=factor(W1empsmum),y=W8QDEB2)) +
  geom_boxplot()
  
W1empsmum
```

#"W1empsdad" vs "W8QDEB2"
```{r}
W1empsdad<-ggplot(data=data, aes(x=factor(W1empsdad),y=W8QDEB2)) +
  geom_boxplot()
  
W1empsdad
```

#"IndSchool" vs "W8QDEB2"
```{r}
IndSchool<-ggplot(data=data, aes(x=factor(IndSchool),y=W8QDEB2)) +
  geom_boxplot()
  
IndSchool
```

#"W1marstatmum" vs "W8QDEB2"
```{r}
W1marstatmum<-ggplot(data=data, aes(x=factor(W1marstatmum),y=W8QDEB2)) +
  geom_boxplot()
  
W1marstatmum
```

#"W1depkids" vs "W8QDEB2"
```{r}
W1depkids<-ggplot(data=data, aes(x=factor(W1depkids),y=W8QDEB2)) +
  geom_boxplot()
  
W1depkids
```
#"W1famtyp2" vs "W8QDEB2"
```{r}
W1famtyp2<-ggplot(data=data, aes(x=factor(W1famtyp2),y=W8QDEB2)) +
  geom_boxplot()
  
W1famtyp2
```

#"W1nssecfam" vs "W8QDEB2"
```{r}
W1nssecfam<-ggplot(data=data, aes(x=factor(W1nssecfam),y=W8QDEB2)) +
  geom_boxplot()
  
W1nssecfam
```

#"W1ethgrpYP" vs "W8QDEB2"
```{r}
W1ethgrpYP<-ggplot(data=data, aes(x=factor(W1ethgrpYP),y=W8QDEB2)) +
  geom_boxplot()
  
W1ethgrpYP
```

#"W1heposs9YP" vs "W8QDEB2"
```{r}
W1heposs9YP<-ggplot(data=data, aes(x=factor(W1heposs9YP),y=W8QDEB2)) +
  geom_boxplot()
  
W1heposs9YP
```

#"W1hwndayYP" vs "W8QDEB2"
```{r}
W1hwndayYP<-ggplot(data=data, aes(x=factor(W1hwndayYP),y=W8QDEB2)) +
  geom_boxplot()
  
W1hwndayYP
```

#"W1truantYP" vs "W8QDEB2"
```{r}
W1truantYP<-ggplot(data=data, aes(x=factor(W1truantYP),y=W8QDEB2)) +
  geom_boxplot()
  
W1truantYP
```

#"W1alceverYP" vs "W8QDEB2"
```{r}
W1alceverYP<-ggplot(data=data, aes(x=factor(W1alceverYP),y=W8QDEB2)) +
  geom_boxplot()
  
W1alceverYP
```

#"W1bulrc" vs "W8QDEB2"
```{r}
W1bulrc<-ggplot(data=data, aes(x=factor(W1bulrc),y=W8QDEB2)) +
  geom_boxplot()
  
W1bulrc
```

#"W1disabYP" vs "W8QDEB2"
```{r}
W1disabYP<-ggplot(data=data, aes(x=factor(W1disabYP),y=W8QDEB2)) +
  geom_boxplot()
  
W1disabYP
```

# W2disc1YP vs W8QDEB2
```{r}
W2disc1YP <- ggplot(data=data, aes(x=factor(W2disc1YP), y=W8QDEB2)) +
  geom_boxplot()
W2disc1YP
```

# W2depressYP vs W8QDEB2
```{r}
W2depressYP <- ggplot(data=data, aes(x=factor(W2depressYP), y=W8QDEB2)) +
  geom_boxplot()
W2depressYP
```

# W4AlcFreqYP vs W8QDEB2
```{r}
W4AlcFreqYP <- ggplot(data=data, aes(x=factor(W4AlcFreqYP), y=W8QDEB2)) + geom_boxplot()
W4AlcFreqYP
```

# W4CannTryYP vs W8QDEB2
```{r}
W4CannTryYP <- ggplot(data=data, aes(x=factor(W4CannTryYP), y=W8QDEB2)) + geom_boxplot()
W4CannTryYP
```

# W4NamesYP vs W8QDEB2
```{r}
W4NamesYP <- ggplot(data=data, aes(x=factor(W4NamesYP), y=W8QDEB2)) + geom_boxplot()
W4NamesYP
```

# W4RacismYP vs W8QDEB2
```{r}
W4RacismYP <- ggplot(data=data, aes(x=factor(W4RacismYP), y=W8QDEB2)) + geom_boxplot()
W4RacismYP
```

# W4empsYP vs W8QDEB2
```{r}
W4empsYP <- ggplot(data=data, aes(x=factor(W4empsYP), y=W8QDEB2)) + geom_boxplot()
W4empsYP
```

# W4Childck1YP vs W8QDEB2
```{r}
W4Childck1YP <- ggplot(data=data, aes(x=factor(W4Childck1YP), y=W8QDEB2)) + geom_boxplot()
W4Childck1YP
```

# W5JobYP vs W8QDEB2
```{r}
W5JobYP <- ggplot(data=data, aes(x=factor(W5JobYP), y=W8QDEB2)) + geom_boxplot()
W5JobYP
```

# W5EducYP vs W8QDEB2
```{r}
W5EducYP <- ggplot(data=data, aes(x=factor(W5EducYP), y=W8QDEB2)) + geom_boxplot()
W5EducYP
```

# W5Apprent1YP vs W8QDEB2
```{r}
W5Apprent1YP <- ggplot(data=data, aes(x=factor(W5Apprent1YP), y=W8QDEB2)) + geom_boxplot()
W5Apprent1YP
```

# W6JobYP vs W8QDEB2
```{r}
W6JobYP <- ggplot(data=data, aes(x=factor(W6JobYP), y=W8QDEB2)) + geom_boxplot()
W6JobYP
```

# W6UnivYP vs W8QDEB2
```{r}
W6UnivYP <- ggplot(data=data, aes(x=factor(W6UnivYP), y=W8QDEB2)) + geom_boxplot()
W6UnivYP
```

# W6EducYP vs W8QDEB2
```{r}
W6EducYP <- ggplot(data=data, aes(x=factor(W6EducYP), y=W8QDEB2)) + geom_boxplot()
W6EducYP
```

# W6Apprent1YP vs W8QDEB2
```{r}
W6Apprent1YP <- ggplot(data=data, aes(x=factor(W6Apprent1YP), y=W8QDEB2)) + geom_boxplot()
W6Apprent1YP
```

# W6acqno vs W8QDEB2
```{r}
W6acqno <- ggplot(data=data, aes(x=factor(W6acqno), y=W8QDEB2)) + geom_boxplot()
W6acqno
```

# W6gcse vs W8QDEB2
```{r}
W6gcse <- ggplot(data=data, aes(x=factor(W6gcse), y=W8QDEB2)) + geom_boxplot()
W6gcse
```

# W6als vs W8QDEB2
```{r}
W6als <- ggplot(data=data, aes(x=factor(W6als), y=W8QDEB2)) + geom_boxplot()
W6als
```

# W6OwnchiDV vs W8QDEB2
```{r}
W6OwnchiDV <- ggplot(data=data, aes(x=factor(W6OwnchiDV), y=W8QDEB2)) + geom_boxplot()
W6OwnchiDV
```

# W6Childliv vs W8QDEB2
```{r}
W6Childliv <- ggplot(data=data, aes(x=factor(W6Childliv), y=W8QDEB2)) + geom_boxplot()
W6Childliv
```

# W8DMARSTAT vs W8QDEB2
```{r}
W8DMARSTAT <- ggplot(data=data, aes(x=factor(W8DMARSTAT), y=W8QDEB2)) + geom_boxplot()
W8DMARSTAT
```

# W8TENURE vs W8QDEB2
```{r}
W8TENURE <- ggplot(data=data, aes(x=factor(W8TENURE), y=W8QDEB2)) + geom_boxplot()
W8TENURE
```

# W8DACTIVITY vs W8QDEB2
```{r}
W8DACTIVITY <- ggplot(data=data, aes(x=factor(W8DACTIVITY), y=W8QDEB2)) + geom_boxplot()
W8DACTIVITY
```

Exploratory analysis of continuous variables

continuous variables: "W1GrssyrHH", "W1yschat1", "W2ghq12scr", "W4schatYP", "W6DebtattYP", "W8DGHQSC"

#"W1GrssyrHH" vs "W8QDEB2"
```{r}
W1GrssyrHH<-ggplot(data, aes(x = W1GrssyrHH, y = W8QDEB2)) + 
  geom_point(alpha=0.5)

W1GrssyrHH
```

#"W1yschat1" vs "W8QDEB2"
```{r}
W1yschat1<-ggplot(data, aes(x = W1yschat1, y = W8QDEB2)) + 
  geom_point(alpha=0.5)

W1yschat1
```

# W2ghq12scr vs W8QDEB2
```{r}
W2ghq12scr <- ggplot(data=data, aes(x=W2ghq12scr, y=W8QDEB2)) +
  geom_point(alpha=0.5)
W2ghq12scr
```

# W4schatYP vs W8QDEB2
```{r}
W4schatYP <- ggplot(data=data, aes(x=W4schatYP, y=W8QDEB2)) + geom_point(alpha=0.5)
W4schatYP
```

# W6DebtattYP vs W8QDEB2
```{r}
W6DebtattYP <- ggplot(data=data, aes(x=W6DebtattYP, y=W8QDEB2)) + geom_point(alpha=0.5)
W6DebtattYP
```

# W8DGHQSC vs W8QDEB2
```{r}
W8DGHQSC <- ggplot(data=data, aes(x=W8DGHQSC, y=W8QDEB2)) + geom_point(alpha=0.5)
W8DGHQSC
```