Setup
```{r}
rm(list=ls())
library(arm)
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
```

# Load data

cleaned data
```{r}
data<-read.csv("../data/W8QDEB2_cleaned.csv",header = TRUE, stringsAsFactors=TRUE)
head(data)
summary(data)
```

# Model 1

run regression on data
```{r}
model.1 <- lm(W8QDEB2 ~ ., data = data)

# have a look at the model
summary(model.1)
```
Keep predictors with more than 5% significant value in the model: "W1GrssyrHH", "IndSchool", "W1depkids", "W4RacismYP"

# Model 2

run regression with predictors having more than 5% significant value from the previous model 

```{r}
model.2 <- lm(W8QDEB2 ~ W1GrssyrHH + IndSchool + W1depkids + W4RacismYP, data = data)

# have a look at the model
summary(model.2)
```

# Model 3

run regression with predictors having more than 5% significant value from the previous model 

```{r}
model.3 <- lm(W8QDEB2 ~ W1GrssyrHH + IndSchool + W4RacismYP, data = data)

# have a look at the model
summary(model.3)
```

check for interactions between the three predictors

