---
title: "data_cleaning"
output: html_document
date: "2024-04-01"
---

```{r}
library(dplyr)       
library(tidyr)      
library(readr)      
library(lubridate)
library(arm)
library(tidyselect)
library(ggplot2)
```


```{r setup, include=FALSE}
data <- read.csv("../W8DINCW.csv")
```


# Summary

- CONTINUOUS variables: *W1GrssyrMP*(44.82), *W1GrssyrHH*(49.3),【to remove these predictors from a final model to see if they change the 
regression substantially, provided of course that they are significant】

W1yschat1(1.86), W2ghq12scr(4.09), W4schatYP(0.63), W6DebtattYP(0),W8DINCW(0), W8DGHQSC(3.7)


- BINARY variables: *W8DDEGP*(38.5% missing)


- CATEGORICAL variables with more than 10% missing value: *W1hiqualdad*(25.4), *W1wrkfulldad*(19.15), *W1empsdad*(19.1),

*W4Childck1YP*(99.5), *W6EducYP*(37.9), *W6Apprent1YP*(37.9), *W6Childliv*(97.4), *W6NEETAct* (93.7), *W8DDEGP*(38.5),

*W6Childliv*(97.4), *W6NEETAct*(93.7)

```{r}
neg_percentages <- data %>%
  summarise_all(~mean(.x < 0, na.rm = TRUE) * 100)

print(neg_percentages)
```
# Data Cleaning

Remove the missing values of variables with less than 10% missing value
```{r}
vars_to_modify <- names(neg_percentages)[neg_percentages < 10]

data.tona <- data %>%
  mutate(across(everything(), ~replace(., . < 0, NA))) %>% #changed the negative values to NA
  filter(!if_any(all_of(vars_to_modify), is.na))#number of rows decreases from 5792 to 3265
```

Factorise categorical variables
```{r}
exclude_vars <- c("W1GrssyrMP", "W1GrssyrHH", "W1yschat1", "W2ghq12scr", "W4schatYP", "W6DebtattYP", "W8DINCW", "W8DGHQSC")
data.tona <- data.tona %>%
  mutate(across(where(is.integer) & !all_of(exclude_vars), factor))
```

## New Level for Missing Values

Create a new level for the missing values of categorical variables
```{r}
add_missing_level <- function(x) {
  if(any(is.na(x))) { #whether there are missing values
    x <- factor(x, levels = c(levels(x), "missing")) #create new level
    x[is.na(x)] <- "missing" #replace NA with new level
  }
  return(x)
}

data.nl <- data.tona %>%
  mutate(across(where(is.factor), ~add_missing_level(.)))
```

# Output1, replace missing values with NA

the missing value of categorical variables with less than 10% missing value are removed

```{r}
write.csv(data.tona, file = "W8DINCW_tona.csv")
```


# Output2, create a new level for the missing values of some categorical variables

the missing value of categorical variables with less than 10% missing value are removed

```{r}
write.csv(data.nl, file = "W8DINCW_newlevel.csv")
```



