---
title: "EDA"
output: html_document
date: "2024-04-02"
---
```{r}
rm(list=ls())
library(dplyr)       
library(tidyr)      
library(readr)      
library(lubridate)
library(arm)
library(tidyselect)
library(ggplot2)
library(gridExtra)
```

```{r}
eot.dat <- read.csv("../data/W8DINCW_newlevel.csv",stringsAsFactors=TRUE)
```

```{r}
data.tona <- read.csv("../data/W8DINCW_tona.csv")
```

# Summary

W8DINCW <- Continuous weekly income according to the survey on the 8th wave

Categorical Variables that are significant to the model(p-value < 0.05):

W1wrk1aMP, W1condur5MP, W1hea2MP, W1NoldBroHS, W1hous12HH, W1usevcHH,

W1hiqualdad, W1hiqualmum, W1wrkfulldad, W1wrkfullmum, W1empsmum, W1empsdad,

W1ch0_2HH, W1ch3_11HH, IndSchool, W1marstatmum, W1depkids, W1famtyp2,

W1nssecfam, W1ethgrpYP, W1heposs9YP(), W1hwndayYP(), W1truantYP, W1alceverYP,

W1bulrc, W1disabYP, W2disc1YP, W2depressYP(), W4AlcFreqYP(only 2 levels are significant), W4CannTryYP,

W4NamesYP(), W4RacismYP, W4empsYP(),W4Childck1YP(overwhelmingly missing data), W5JobYP, W5JobYP, 

W5EducYP, W5Apprent1YP, W6JobYP, W6UnivYP, W6EducYP, W6Apprent1YP(),

W6acqno, W6gcse(), W6als(more significant than w6gcse), W6OwnchiDV, W6Childliv, W6NEETAct(the only significant level is missing level),

W8DDEGP, W8TENURE


Continuous Varaibles that are significant to the model(p-value < 0.05):

W2ghq12scr(), W4schatYP(), W6DebtattYP



May wanna merge some levels of: W1hiqualmum/dad, W4AlcFreqYP

# Categorical variables

## W1wrk1aMP vs W8DINCW

Main Parents' current working status

Value = 1.0 Label = Full-time paid employee (30 or more hours a week)
Value = 2.0 Label = Part-time paid employee (under 30 hours a week)
Value = 3.0 Label = Full-time self-employed
Value = 4.0 Label = Part-time self-employed
Value = 5.0 Label = Unemployed and seeking work
Value = 6.0 Label = Full-time education
Value = 7.0 Label = On a government scheme for employment training
Value = 8.0 Label = Temporarily sick/ disabled (if temporarily off sick from job
Value = 9.0 Label = Permanently sick/ disabled
Value = 10.0 Label = Looking after home/ family
Value = 11.0 Label = Retired from work altogether
Value = 12.0 Label = Other answers

```{r}
ggplot(data=eot.dat, aes(x=factor(W1wrk1aMP),y=W8DINCW)) +
  geom_boxplot()
```
It is clear that the median weekly income of parents who are currently employed(1-4) is higher than the rest of the categories, and the median weekly income of parents who are unemployed(5) and seeking work is the lowest.

```{r}
summary(lm(W8DINCW ~ factor(W1wrk1aMP), data=eot.dat))
```
It shows that the p-value is less than 0.05, which means that the working status of the main parent is significant to the weekly income.

Full-time self-employed(3) and On a government scheme for employment training(7) are not significant. Maybe due to the small sample size.

## W1condur5MP vs W8DINCW

Whether have home computer in household.
1-yes, 2-no

```{r}
ggplot(data=eot.dat, aes(x=factor(W1condur5MP),y=W8DINCW)) +
  geom_boxplot()
```
Household with a home computer has a higher median weekly income than the household without a home computer.

```{r}
summary(lm(W8DINCW ~ factor(W1condur5MP), data=eot.dat))
```
The p-value is less than 0.05, which means that the presence of a home computer in the household is significant to the weekly income.

## W1hea2MP vs W8DINCW

Whether MP has longstanding illness or disability

1-Yes, 2-No

```{r}
ggplot(data=eot.dat, aes(x=factor(W1hea2MP),y=W8DINCW)) +
  geom_boxplot()
```
Intuitively,parents who have a longstanding illness or disability earns less than those who don't.

```{r}
summary(lm(W8DINCW ~ factor(W1hea2MP), data=eot.dat))
```

The p-value is less than 0.05, which means that the presence of a longstanding illness or disability is significant to the weekly income.

## W1NoldBroHS vs W8DINCW

Number of younger siblings

```{r}
ggplot(data=eot.dat, aes(x=factor(W1NoldBroHS),y=W8DINCW)) +
  geom_boxplot()
```
Though similar to 0 and 2, the median weekly income of parents with 1 younger sibling is the highest among all categories. 
There is a decreasing pattern in the median weekly income as the number of younger siblings increases from 1 to 4. The median weekly income of parents with 5 or more younger siblings is the lowest while the data is the most dispersed.

```{r}
summary(lm(W8DINCW ~ factor(W1NoldBroHS), data=eot.dat))
```
The p-value is less than 0.05, with the categories of having 4 or 5 siblings shows relatively great significance to the weekly income.

## W1InCarHH vs W8DINCW
Whether YP(young people, the participants) is living in a private HH(household) or not

Value = 1.0 Label = Private household
Value = 2.0 Label = Hotel/ B&B
Value = 3.0 Label = Institution
Value = 4.0 Label = Other answers
```{r}
ggplot(data=eot.dat, aes(x=factor(W1InCarHH),y=W8DINCW)) +
  geom_boxplot()

#Before data cleaning, this variable looks like:
#-999    1    2    3    4 
#  25 5752    6    1    8

#After data cleaning, this variable looks like:
#   1    4 
#3260    5
```

The median weekly income of parents who are living in a private household is higher than those who are living in a hotel/B&B.

```{r}
summary(lm(W8DINCW ~ factor(W1InCarHH), data=eot.dat))
```
This variable is not significant to the weekly income. This could be due to the small sample size of the categories other than 1.

## W1hous12HH vs W8DINCW

Tenure

Value = 1.0 Label = Owned outright
Value = 2.0 Label = Being bought on a mortgage/ bank loan
Value = 3.0 Label = Shared ownership (owns & rents property)
Value = 4.0 Label = Rented from a Council or New Town
Value = 5.0 Label = Rented from a Housing Association
Value = 6.0 Label = Rented privately
Value = 7.0 Label = Rent free
Value = 8.0 Label = Some other arrangement

```{r}
ggplot(data=eot.dat, aes(x=factor(W1hous12HH),y=W8DINCW)) +
  geom_boxplot()
```
The median weekly income of people who owns and rents the property(3) has the highest median weekly income, followed by people who has some other arrangement(6). The median weekly income of people who rent from a council or new town(4) and Housing Association(5) is the lowest.

People whose household is rent free(2) has similar income with who own the house outright(1).


```{r}
summary(lm(W8DINCW ~ factor(W1hous12HH), data=eot.dat))

# 1    2     3    4    5    6     7   8 
#435  2226   11  290  182  100    4   17

#The insignificant p-value of the categories 3, 6, 7, 8 could be due to the small sample size. While 3(Shared ownership (owns & rents property)) shows a great significance to the weekly income.
```

## W1usevcHH vs W8DINCW

Whether anyone uses a motor vehicle (1-yes, 2-no)

```{r}
ggplot(data=eot.dat, aes(x=factor(W1usevcHH),y=W8DINCW)) +
  geom_boxplot()
```
The median weekly income of people who use a motor vehicle is higher than those who don't.

```{r}
summary(lm(W8DINCW ~ factor(W1usevcHH), data=eot.dat))
```

## W1hiqualmum vs W8DINCW & W1hiqualdad vs W8DINCW

Mother's highest qualification and Father's highest qualification

Value = 1.0 Label = Higher Degree -- postgraduare,phd..

Value = 2.0 Label = First Degree -- bachelor

Value = 3.0 Label = HE Diploma -- between A level and bachelor
Value = 4.0 Label = HNC/HND/NVQ4 -- between GCSE and A level, but better
Value = 5.0 Label = Teaching qualification, non-degree
Value = 6.0 Label = Nursing qualification, non-degree
Value = 7.0 Label = A Levels
Value = 8.0 Label = OND/ONC -- between GCSE and A level

Value = 9.0 Label = City and guilds part III, NVQ3 -- between GCSE and A level
Value = 10.0 Label = CSYS -- between GCSE and A level
Value = 11.0 Label = Scottish Higher Grade -- A level
Value = 12.0 Label = AS Level
Value = 13.0 Label = Trade apprenticeship
Value = 14.0 Label = City and guilds part II, NVQ2

Value = 15.0 Label = GCSE grade A-C and equivalent

Value = 16.0 Label = GCSE grade D-E and equivalent
Value = 17.0 Label = City and guilds part I, NVQ1
Value = 18.0 Label = Youth training, skill seekers
Value = 19.0 Label = Qualification, level unspecified
Value = 20.0 Label = No qualification mentioned

```{r}
mhq_g <- ggplot(data=eot.dat, aes(x=factor(W1hiqualmum),y=W8DINCW)) +
  geom_boxplot()

dhq_g <- ggplot(data=eot.dat, aes(x=factor(W1hiqualdad),y=W8DINCW)) +
  geom_boxplot()

grid.arrange(mhq_g, dhq_g, nrow=2)
```

```{r}
summary(lm(W8DINCW ~ factor(W1hiqualmum), data=eot.dat))
```

```{r}
summary(lm(W8DINCW ~ factor(W1hiqualdad), data=eot.dat))
```

## W1wrkfulldad vs W8DINCW & W1wrkfullmum vs W8DINCW

Whether father/mother works full-time

Value = 1.0 Label = Working full-time
Value = 2.0 Label = Working part-time
Value = 3.0 Label = Currently not working

```{r}
wfd_g <- ggplot(data=eot.dat, aes(x=factor(W1wrkfulldad),y=W8DINCW)) +
  geom_boxplot()

wfm_g <- ggplot(data=eot.dat, aes(x=factor(W1wrkfullmum),y=W8DINCW)) +
  geom_boxplot()

grid.arrange(wfd_g, wfm_g, nrow=2)
```

The weekly income of young people whose father work full-time is higher than those who work part-time or not working.

The weekly income of young people whose mother work part-time is higher than those who work full-time or not working.

```{r}
summary(lm(W8DINCW ~ factor(W1wrkfulldad), data=eot.dat))
```

```{r}
summary(lm(W8DINCW ~ factor(W1wrkfullmum), data=eot.dat))
```
## W1empsmum vs W8DINCW & W1empsdad vs W8DINCW

Employment status of mother/father

Value = 1.0 Label = Doing paid work for 30 or more hours a week
Value = 2.0 Label = Doing paid work for fewer than 30 hours a week
Value = 3.0 Label = Unemployed/ Looking for a job
Value = 4.0 Label = On a training course or scheme
Value = 5.0 Label = In full-time education/ at school
Value = 6.0 Label = Looking after the family/ household
Value = 7.0 Label = Retired from work altogether
Value = 8.0 Label = Sick/ disabled
Value = 9.0 Label = Other

```{r}
em_g <- ggplot(data=eot.dat, aes(x=factor(W1empsmum),y=W8DINCW)) +
  geom_boxplot()

ed_g <- ggplot(data=eot.dat, aes(x=factor(W1empsdad),y=W8DINCW)) +
  geom_boxplot()

grid.arrange(em_g, ed_g, nrow=2)
```
The weekly income of young people whose mother is doing paid work fewer than 30 hours a week is the highest among all categories. 

This corresponds to the info that the weekly income of young people whose mother is doing part-time job is higher than those who are doing full-time job.

The weekly income of young people whose father is doing paid work for 30 or more hours a week is the highest among all categories.

```{r}
summary(lm(W8DINCW ~ factor(W1empsmum), data=eot.dat))
```

```{r}
summary(lm(W8DINCW ~ factor(W1empsdad), data=eot.dat))
```

## W1ch0_2HH,W1ch3_11HH,W1ch12_15HH,W1ch16_17HH vs W8DINCW

Number of children aged 0-2, 3-11, 12-15, 16-17 in the household

```{r}
a02g <-ggplot(data=eot.dat, aes(x=factor(W1ch0_2HH),y=W8DINCW)) +
  geom_boxplot()
a311g <-ggplot(data=eot.dat, aes(x=factor(W1ch3_11HH),y=W8DINCW)) +
  geom_boxplot()
a1215g <-ggplot(data=eot.dat, aes(x=factor(W1ch12_15HH),y=W8DINCW)) +
  geom_boxplot()
a1617g <-ggplot(data=eot.dat, aes(x=factor(W1ch16_17HH),y=W8DINCW)) +
  geom_boxplot()

grid.arrange(a02g, a311g, a1215g, a1617g, nrow=2)
```
Less number of children in any age group in the household, higher the median weekly income.

```{r}
summary(lm(W8DINCW ~ factor(W1ch0_2HH)+factor(W1ch3_11HH)+factor(W1ch12_15HH)+factor(W1ch16_17HH), data=eot.dat))
```
W1ch0_2HH and W1ch3_11HH are significant to the weekly income.

## IndSchool vs W8DINCW

Whether young people was at an independent or maintained school at sampling stage (1=Independent, 0=Maintained)

```{r}
ggplot(data=eot.dat, aes(x=factor(IndSchool),y=W8DINCW)) +
  geom_boxplot()
```
The median weekly income of young people who was at an independent school is higher than those who was at a maintained school.

```{r}
summary(lm(W8DINCW ~ factor(IndSchool), data=eot.dat))
```
Significant to the weekly income.

## W1marstatmum vs W8DINCW

Marital status of young person's mother

Value = 1.0 Label = Single, that is, never married
Value = 2.0 Label = Married and living with husband or wife
Value = 3.0 Label = Living with a partner
Value = 4.0 Label = Married and separated from husband or wife
Value = 5.0 Label = Divorced
Value = 6.0 Label = Widowed
Value = 7.0 Label = Other

```{r}
ggplot(data=eot.dat, aes(x=factor(W1marstatmum),y=W8DINCW)) +
  geom_boxplot()
```
The median weekly income of young people whose mother is married and living with husband or wife is the highest among all categories.

```{r}
summary(lm(W8DINCW ~ factor(W1marstatmum), data=eot.dat))
```

## W1depkids vs W8DINCW

Number of dependent children in Household

```{r}
ggplot(data=eot.dat, aes(x=factor(W1depkids),y=W8DINCW)) +
  geom_boxplot()
```
The median weekly income of young people whose household has 2 dependent child is the highest among all categories. It shows similar pattern with the number of children in the household.


```{r}
summary(lm(W8DINCW ~ factor(W1depkids), data=eot.dat))
```

## W1famtyp2 vs W8DINCW

Whether single parent family or not(0-no,1-yes)

```{r}
ggplot(data=eot.dat, aes(x=factor(W1famtyp2),y=W8DINCW)) +
  geom_boxplot()
```
The median weekly income of young people who are from a single parent family is lower than those who are not.

```{r}
summary(lm(W8DINCW ~ factor(W1famtyp2), data=eot.dat))
```
## W1nssecfam vs W8DINCW

Family's NS_SEC class

Value = 1.0 Label = Higher Managerial and professional occupations
Value = 2.0 Label = Lower managerial and professional occupations
Value = 3.0 Label = Intermediate occupations
Value = 4.0 Label = Small employers and own account workers
Value = 5.0 Label = Lower supervisory and technical occupations
Value = 6.0 Label = Semi-routine occupations
Value = 7.0 Label = Routine occupations
Value = 8.0 Label = Never worked/long term unemployed

```{r}
ggplot(data=eot.dat, aes(x=factor(W1nssecfam),y=W8DINCW)) +
  geom_boxplot()
```

The median weekly income of young people whose family's NS_SEC class is higher managerial and professional occupations(1) is the highest among all categories.

```{r}
summary(lm(W8DINCW ~ factor(W1nssecfam), data=eot.dat))
```

## W1ethgrpYP vs W8DINC

Young person's ethnic group

Value = 1.0 Label = White
Value = 2.0 Label = Mixed
Value = 3.0 Label = Indian
Value = 4.0 Label = Pakistani
Value = 5.0 Label = Bangladeshi
Value = 6.0 Label = Black Caribbean
Value = 7.0 Label = Black African
Value = 8.0 Label = Other

```{r}
ggplot(data=eot.dat, aes(x=factor(W1ethgrpYP),y=W8DINCW)) +
  geom_boxplot()
```

White people have the highest weekly income among all categories.


```{r}
summary(lm(W8DINCW ~ factor(W1ethgrpYP), data=eot.dat))
```

## W1heposs9YP vs W8DINCW

Likelihood of Young Person going to University

Value = 1.0 Label = Very likely
Value = 2.0 Label = Fairly likely
Value = 3.0 Label = Not very likely
Value = 4.0 Label = Not at all likely

```{r}
ggplot(data=eot.dat, aes(x=factor(W1heposs9YP),y=W8DINCW)) +
  geom_boxplot()
```

Higher the likelihood of going to university, higher the median weekly income.


```{r}
summary(lm(W8DINCW ~ factor(W1heposs9YP), data=eot.dat))
```

## W1hwndayYP vs W8DINCW

Number of evenings do homework per week

```{r}
ggplot(data=eot.dat, aes(x=factor(W1hwndayYP),y=W8DINCW)) +
  geom_boxplot()
```

People who spend 4 evenings doing homework per week have the highest median weekly income.


```{r}
summary(lm(W8DINCW ~ factor(W1hwndayYP), data=eot.dat))
```

## W1truantYP vs W8DINCW

Whether YP played truant in last 12 months

```{r}
ggplot(data=eot.dat, aes(x=factor(W1truantYP),y=W8DINCW)) +
  geom_boxplot()
```

```{r}
summary(lm(W8DINCW ~ factor(W1truantYP), data=eot.dat))
```

## W1alceverYP vs W8DINCW

Whether YP has ever been drunk

```{r}
ggplot(data=eot.dat, aes(x=factor(W1alceverYP),y=W8DINCW)) +
  geom_boxplot()
```

```{r}
summary(lm(W8DINCW ~ factor(W1alceverYP), data=eot.dat))

```
## W1bulrc vs W8DINCW

Whether YP has been bullied in the last 12 months

```{r}
ggplot(data=eot.dat, aes(x=factor(W1bulrc),y=W8DINCW)) +
  geom_boxplot()
```

```{r}
summary(lm(W8DINCW ~ factor(W1bulrc), data=eot.dat))
```

## W1disabYP vs W8DINCW

Whether YP has a disability/long term illness

```{r}
ggplot(data=eot.dat, aes(x=factor(W1disabYP),y=W8DINCW)) +
  geom_boxplot()
```

```{r}
summary(lm(W8DINCW ~ factor(W1disabYP), data=eot.dat))

```

## W2disc1YP vs W8DINCW

Whether YP thinks they have ever been treated unfairly by teachers because of skin colour or ethnic origin

1-yes, 2-no

```{r}
ggplot(data=eot.dat, aes(x=factor(W2disc1YP),y=W8DINCW)) +
  geom_boxplot()
```

The median weekly income of young people who thinks they have ever been treated unfairly by teachers... is higher than those who don't.


```{r}
summary(lm(W8DINCW ~ factor(W2disc1YP), data=eot.dat))
```

## W2depressYP vs W8DINCW

Whether YP has been feeling unhappy or depressed recently

Value = 1.0 Label = Not at all
Value = 2.0 Label = No more than usual
Value = 3.0 Label = Rather more than usual
Value = 4.0 Label = Much more than usual

```{r}
ggplot(data=eot.dat, aes(x=factor(W2depressYP),y=W8DINCW)) +
  geom_boxplot()
```

The weekly income of young people who feels depressed relatively more often is higher than those who don't.


```{r}
summary(lm(W8DINCW ~ factor(W2depressYP), data=eot.dat))
```
## W4AlcFreqYP vs W8DINCW

Frequency of having an alcoholic drink in last 12 months
 
Value = 1.0 Label = Most days
Value = 2.0 Label = Once or twice a week
Value = 3.0 Label = 2 or 3 times a month
Value = 4.0 Label = Once a month
Value = 5.0 Label = Once every couple of months or
Value = 6.0 Label = Less often?

```{r}
ggplot(data=eot.dat, aes(x=factor(W4AlcFreqYP),y=W8DINCW)) +
  geom_boxplot()
```

The median weekly income of people who drunk more than once a month is similar.
The“Once every few months” and“Less” frequencies of income distribution seem to have more high-income outliers, indicating that at these frequencies, a small number of people earn significantly more per week than others

```{r}
summary(lm(W8DINCW ~ factor(W4AlcFreqYP), data=eot.dat))
```
## W4CannTryYP vs W8DINCW

Whether YP ever tried cannabis

1-yes, 2-no

```{r}
ggplot(data=eot.dat, aes(x=factor(W4CannTryYP),y=W8DINCW)) +
  geom_boxplot()
```

The median weekly income of people who have tried cannabis is higher than those who haven't.

```{r}
summary(lm(W8DINCW ~ factor(W4CannTryYP), data=eot.dat))
```

## W4NamesYP vs W8DINCW

Whether YP has been called names, sworn at or insulted in last 12 months

```{r}
ggplot(data=eot.dat, aes(x=factor(W4NamesYP),y=W8DINCW)) +
  geom_boxplot()
```

The median weekly income of people who have been called names, sworn at or insulted in the last 12 months is higher than those who haven't.

```{r}
summary(lm(W8DINCW ~ factor(W4NamesYP), data=eot.dat))
```

## W4RacismYP vs W8DINCW

Whether YP has been threatened/insulted in the last 12 months due to skin colour/ethnicity

```{r}
ggplot(data=eot.dat, aes(x=factor(W4RacismYP),y=W8DINCW)) +
  geom_boxplot()

```

People who haven't been threatened/insulted due to skin colour has significantly higher weekly income than those who have.

```{r}
summary(lm(W8DINCW ~ factor(W4RacismYP), data=eot.dat))
```

## W4empsYP vs W8DINCW

Employment status of young person

Value = 1.0 Label = Doing paid work for 30 or more hours a week
Value = 2.0 Label = Doing paid work for fewer than 30 hours a week
Value = 3.0 Label = Unemployed/ Looking for a job
Value = 4.0 Label = On a training course or scheme
Value = 5.0 Label = In full-time education/ at school
Value = 6.0 Label = Looking after the family/ household
Value = 7.0 Label = Retired from work altogether
Value = 8.0 Label = Sick/ disabled
Value = 9.0 Label = Other

```{r}
ggplot(data=eot.dat, aes(x=factor(W4empsYP),y=W8DINCW)) +
  geom_boxplot()

```

Young person with different employment status has different weekly income distribution, where the median weekly income of people who are doing paid work for 30 or more hours a week is the highest among all categories.

```{r}
summary(lm(W8DINCW ~ factor(W4empsYP), data=eot.dat))
```

## W4Childck1YP vs W8DINCW

Interviewer confirm Young People has child

1-yes, 2-no

```{r}
ggplot(data=eot.dat, aes(x=factor(W4Childck1YP),y=W8DINCW)) +
  geom_boxplot()

#before data cleaning  -- -99, not interviewd.  -91, not applicable
# -99   -91     1 
#  37   5728    27

#After data cleaning
# 1     missing 
# 12    3253
```

The median weekly income of people who have children is higher than those who don't.

```{r}
summary(lm(W8DINCW ~ factor(W4Childck1YP), data=eot.dat))
```

## W5JobYP, W5EducYP, W5Apprent1YP vs W8DINCW

Young Person: Whether currently doing paid work at 5th wave
YP: Whether currently going to school or college at 5th wave
YP: Whether currently doing an apprenticeship at 5th wave

1-yes, 2-no

```{r}
w5j_g <- ggplot(data=eot.dat, aes(x=factor(W5JobYP),y=W8DINCW)) +
  geom_boxplot()
w5e_g <- ggplot(data=eot.dat, aes(x=factor(W5EducYP),y=W8DINCW)) +
  geom_boxplot()
w5a_g <- ggplot(data=eot.dat, aes(x=factor(W5Apprent1YP),y=W8DINCW)) +
  geom_boxplot()

grid.arrange(w5j_g, w5e_g, w5a_g, nrow=2)
```

The median weekly income are higher for people who are doing paid work, going to school or colledge, and who are not doing an apprenticeship at 5th wave.

```{r}
summary(lm(W8DINCW ~ factor(W5JobYP)+, data=eot.dat))
```
```{r}
summary(lm(W8DINCW ~ factor(W5EducYP), data=eot.dat))
```

```{r}
summary(lm(W8DINCW ~ factor(W5Apprent1YP), data=eot.dat))
```

## W6JobYP, W6UnivYP, W6EducYP, W6Apprent1YP vs W8DINCW

YP: Whether currently doing paid work at 6th wave
YP: Whether currently at university at 6th wave
YP: Whether currently going to school or college at 6th wave
YP: Whether currently doing an Apprenticeship at 6th wave

```{r}
w6j_g <- ggplot(data=eot.dat, aes(x=factor(W6JobYP),y=W8DINCW)) +
  geom_boxplot()
w6u_g <- ggplot(data=eot.dat, aes(x=factor(W6UnivYP),y=W8DINCW)) +
  geom_boxplot()
w6e_g <- ggplot(data=eot.dat, aes(x=factor(W6EducYP),y=W8DINCW)) +
  geom_boxplot()
w6a_g <- ggplot(data=eot.dat, aes(x=factor(W6Apprent1YP),y=W8DINCW)) +
  geom_boxplot()

grid.arrange(w6j_g, w6u_g, w6e_g, w6a_g, nrow=2)
```

Similar to the 5th wave, the median weekly income are higher for people who are doing paid work, and who are not doing an apprenticeship at 6th wave.

However, the median weekly income of people who going to school or college is lower than those who don't at 6th wave.

The weekly income is higher for people who are at university is higher than those who are not. The increase of this sub group coould explain the different pattern of W6EducYP and 56EducYP.

The median weekly income is higher for the missing category of W6EducYP and W6Apprent1YP.

```{r}
summary(lm(W8DINCW ~ factor(W6JobYP), data=eot.dat))
```

```{r}
summary(lm(W8DINCW ~ factor(W6UnivYP), data=eot.dat))
```

```{r}
summary(lm(W8DINCW ~ factor(W6EducYP), data=eot.dat))
```

```{r}
summary(lm(W8DINCW ~ factor(W6Apprent1YP), data=eot.dat))
```

## W6acqno vs W8DINCW

Highest academic qualification studied at Wave 6

alue = 1.0 Label = First/Other Degree
Value = 2.0 Label = Other HE
Value = 3.0 Label = 2+ A/AS
Value = 4.0 Label = 1 A/AS
Value = 5.0 Label = 5+ GCSE
Value = 6.0 Label = 1-4 GCSE
Value = 7.0 Label = Other
Value = 8.0 Label = Insufficient detail
Value = 9.0 Label = No academic stud

```{r}
ggplot(data=eot.dat, aes(x=factor(W6acqno),y=W8DINCW)) +
  geom_boxplot()
```

Young person with different highest academic qualification studied at Wave 6 has different weekly income distribution, where the weekly income of people who has higher academic qualification(1&2) is the highest among all categories.

People with 5+ GCSE(5) has the lowest median weekly income.

```{r}
summary(lm(W8DINCW ~ factor(W6acqno), data=eot.dat))
```

## W6gcse, W6als vs W8DINCW

Number of GCSEs and A levels  being studied at Wave 6

```{r}
gcse_g <- ggplot(data=eot.dat, aes(x=factor(W6gcse),y=W8DINCW)) +
  geom_boxplot()
als_g <- ggplot(data=eot.dat, aes(x=factor(W6als),y=W8DINCW)) +
  geom_boxplot()
  
grid.arrange(gcse_g, als_g, nrow=2)
```

The median weekly income of people who being studied 3 GCSE or A-levels are the lowest. Those who are studying 4 GCSE or A-levels have the highest median weekly income.

```{r}
summary(lm(W8DINCW ~ factor(W6gcse), data=eot.dat))
```

```{r}
summary(lm(W8DINCW ~ factor(W6als), data=eot.dat))
```

## W6OwnchiDV vs W8DINCW

Whether Respondents have own Child/Children

```{r}
ggplot(data=eot.dat, aes(x=factor(W6OwnchiDV),y=W8DINCW)) +
  geom_boxplot()
```

The median weekly income of people who do not own child/children is higher than those who do.

```{r}
summary(lm(W8DINCW ~ factor(W6OwnchiDV), data=eot.dat))
```

## W6Childliv vs W8DINCW

Whether has own Child/Children Living with Them

```{r}
ggplot(data=eot.dat, aes(x=factor(W6Childliv),y=W8DINCW)) +
  geom_boxplot()
```

The median weekly income of people who has own child/children living with them is higher than those who do. 
Missing category has the highest median weekly income.

```{r}
summary(lm(W8DINCW ~ factor(W6Childliv), data=eot.dat))
```


## W6NEETAct vs W8DINCW

Activities of NEETs (Not in Employment or Education)

Value = 1.0 Label = Looking for paid work
Value = 2.0 Label = Looking after family/home full-time
Value = 3.0 Label = Something else

```{r}
ggplot(data=eot.dat, aes(x=factor(W6NEETAct),y=W8DINCW)) +
  geom_boxplot()
```

```{r}
summary(lm(W8DINCW ~ factor(W6NEETAct), data=eot.dat))
```

## W8DDEGP vs W8DINCW

Whether achieved first degree or higher

0-no, 1-yes

```{r}
ggplot(data=eot.dat, aes(x=factor(W8DDEGP),y=W8DINCW)) +
  geom_boxplot()
```

The median weekly income of people who achieved first degree or higher is higher than those who don't.

```{r}
summary(lm(W8DINCW ~ factor(W8DDEGP), data=eot.dat))
```

## W8DMARSTAT vs W8DINCW

Marital status

Value = 1.0 Label = Single and never married or in a CP
Value = 2.0 Label = Married
Value = 3.0 Label = Separated but still legally married
Value = 4.0 Label = Divorced
Value = 5.0 Label = Widowed
Value = 6.0 Label = A Civil Partner
Value = 7.0 Label = Separated but still legally in a CP
Value = 8.0 Label = A former Civil Partner
Value = 9.0 Label = A surviving Civil Partner

```{r}
ggplot(data=eot.dat, aes(x=factor(W8DMARSTAT),y=W8DINCW)) +
  geom_boxplot()
```


```{r}
summary(lm(W8DINCW ~ factor(W8DMARSTAT), data=eot.dat))
```
## W8TENURE vs W8DINCW

Value = 1.0 Label = Own – outright
Value = 2.0 Label = Own – buying with help of mortg or loan
Value = 3.0 Label = Part rent/mortgage (shared/equity own)
Value = 4.0 Label = Rent inc Housing Ben/Lcl Housing Allow
Value = 5.0 Label = Rent-free inc rels/friend exc squatting
Value = 6.0 Label = Squatting
Value = 7.0 Label = Other arrangement inc board to parents

```{r}
ggplot(data=eot.dat, aes(x=factor(W8TENURE),y=W8DINCW)) +
  geom_boxplot()
```

```{r}
summary(lm(W8DINCW ~ factor(W8TENURE), data=eot.dat))

```


# Continuousn variables

## W1yschat1 vs W8DINCW

Young Person's attitude to school

```{r}
attsch_g<-ggplot(eot.dat, aes(x = W1yschat1, y = W8DINCW)) + 
  geom_point(alpha=0.2, size=3, color="lightblue", stroke=1) + 
  geom_smooth(method = "lm",  
              se = FALSE) + 
  theme_minimal() + 
  labs(title = "Scatter Plot of attitude to school vs weekly income with Fitted Line",
       x = "attitude to school",
       y = "Cotinuous weekly income")

attsch_g
```

The vertical spread of data points at each level of the attitude to school suggests that there is variability in weekly income that is not explained by the attitude to school alone. There may be other factors at play affecting weekly income that are not captured in this single-variable analysis

```{r}
summary(lm(W8DINCW ~ W1yschat1, data=eot.dat))
```

## W2ghq12scr vs W8DINCW

GHQ12 score

```{r}
ghq_g <- ggplot(eot.dat, aes(x = W2ghq12scr, y = W8DINCW)) + 
  geom_point(alpha=0.2, size=3, color="lightblue", stroke=1) + 
  geom_smooth(method = "lm",  
              se = FALSE) + 
  theme_minimal() + 
  labs(title = "Scatter Plot of GHQ12 score vs weekly income with Fitted Line",
       x = "GHQ12 score",
       y = "Cotinuous weekly income")

ghq_g
```

It shows that the weekly income is not significantly related to the GHQ12 score.

```{r}
summary(lm(W8DINCW ~ W2ghq12scr, data=eot.dat))
```
## W4schatYP vs W8DINCW

DV: Young Person's attitude to school

```{r}
schat_g <- ggplot(eot.dat, aes(x = W4schatYP, y = W8DINCW)) + 
  geom_point(alpha=0.2, size=3, color="lightblue", stroke=1) + 
  geom_smooth(method = "lm",  
              se = FALSE) + 
  theme_minimal() + 
  labs(title = "Scatter Plot of attitude to school vs weekly income with Fitted Line",
       x = "attitude to school",
       y = "Cotinuous weekly income")

schat_g
```

```{r}
summary(lm(W8DINCW ~ W4schatYP, data=eot.dat))
```

## W6DebtattYP vs W8DINCW

```{r}
debtatt_g <- ggplot(eot.dat, aes(x = W6DebtattYP, y = W8DINCW)) + 
  geom_point(alpha=0.2, size=3, color="lightblue", stroke=1) + 
  geom_smooth(method = "lm",  
              se = FALSE) + 
  theme_minimal() + 
  labs(title = "Scatter Plot of Debt attitude vs weekly income with Fitted Line",
       x = "Debt attitude",
       y = "Cotinuous weekly income")

debtatt_g
```

```{r}
summary(lm(W8DINCW ~ W6DebtattYP, data=eot.dat))
```


## W8DGHQSC vs W8DINCW

General Health Questionnaire (GHQ12) score 

```{r}
ghqscr_g <- ggplot(eot.dat, aes(x = W8DGHQSC, y = W8DINCW)) + 
  geom_point(alpha=0.2, size=3, color="lightblue", stroke=1) + 
  geom_smooth(method = "lm",  
              se = FALSE) + 
  theme_minimal() + 
  labs(title = "Scatter Plot of GHQ12 score vs weekly income with Fitted Line",
       x = "GHQ12 score",
       y = "Cotinuous weekly income")

ghqscr_g
```
```{r}
summary(lm(W8DINCW ~ W8DGHQSC, data=eot.dat))

```

